### 2015-10-12 thesis-proposal/problem.tex
minor corrections

%Well-designed general-purpose programming languages give library providers the power to  modularly express a wide variety of useful constructs using a comparatively small set of primitives. %New primitives, and thus new dialects of these languages, should arise very rarely. 
%New primitives (and new \emph{dialects} of the language). %Extending a general-pandurpose language with new primitives, i.e.  forming an extended \emph{dialect} of the language, should, given a sufficiently expressive such language, be exceedingly rare. %Indeed, a diversity of dialects around a language can be taken as evidence against claims of  generality.%, this can be taken as evidence against claims of its generality. %Not all useful abstractions can be  realized, when considered comprehensively, as libraries.% The problem of achieving both simplicity and generality simultaneously with a single system remains unsolved. 
%Many languages claim to achieve this ideal. 
%For example, 
%If we measure the generality of a language by how frequently  dialects of the language emerge, 
%all major contemporary languages seem to leave room for improvement. 
%General-purpose programming languages are a dime a. 
%Among the chief goals in the design of a general-purpose programming languages must achieve \emph{stability}: they  very rarely needs to be extended or forked into dialects by the user community surrounding it. Instead, its users are able to modularly  express a wide variety of useful constructs in libraries, using the comparatively small set of primitives that the language builds in. 
Functional programming language designers have long studied small typed lambda calculi to develop a principled understanding of language-theoretic issues, like type safety, and to examine the mathematical character of language primitives of possible interest in isolation. These studies have then informed the design of ``full-scale''\footnote{Throughout this work, words and phrases that should be read as having an intuitive or informal meaning, rather than a strict mathematical meaning, will be introduced with quotation marks.} functional languages, which combine several such primitives and also introduce various generalizations and primitive ``embellishments''  motivated by a consideration of human factors. 
For example, major functional languages like Standard ML (SML) \cite{mthm97-for-dart,harper1997programming}, OCaml \cite{ocaml-manual} and Haskell \cite{jones2003haskell} all primitively build in record types, generalizing the nullary and binary product types that suffice in simpler calculi, because explicitly labeled components are cognitively useful to human programmers. Similarly, these languages build in derived syntactic forms (colloquially, ``syntactic sugar'') that decrease the syntactic cost of working with common library constructs, like lists.

@@

Some of these dialects extend the type structure of the language that they are based on. However, the more common situation, and the one that we will be our focus in this work, appears to be that the existing semantic primitives suffice and it is only the \emph{syntactic cost} of expressing one or more constructs of interest using these primitives  that is too high. In response, library providers construct \emph{syntactic dialects} -- dialects that introduce only new derived syntactic forms. 
%Put another way, syntactic dialects can be specified by a context-independent expansion to the existing language that they are based on. 
For example, Ur/Web is a syntactic dialect of Ur (a language that itself descends from ML \cite{conf/pldi/Chlipala10}) that builds in derived forms for SQL queries, HTML elements and other datatypes used in the domain of web programming \cite{conf/popl/Chlipala15}. %Syntactic cost is often assessed qualitatively \cite{green1996usability}, though quantitative metrics can be defined. 
This is not an isolated example -- there are many other types of data that could similarly benefit from the availability of specialized derived forms (we will consider another example, regular expression patterns, in Sec. \ref{sec:syntax-examples}). 
Tools like Camlp4 \cite{ocaml-manual}, Sugar* \cite{erdweg2011sugarj,erdweg2013framework} and Racket \cite{Flatt:2012:CLR:2063176.2063195}, which we will discuss in Sec. \ref{sec:syntax-existing-approaches}, have lowered the engineering costs of constructing syntactic dialects in such situations, further contributing to their proliferation.

@@

\inferrule[T-U-abs]{
	\Delta \vdash \tau_1~\mathtt{type}\\
	\Delta~\Gamma, x:\tau \vdash \hat{e} \leadsto e : \tau'
}{
	\Delta~\Gamma \vdash \lambda x{:}\tau.\hat{e} \leadsto \lambda x{:}\tau.e : \tau \rightharpoonup \tau'hh
}

>>>>

%Well-designed general-purpose programming languages give library providers the power to  modularly express a wide variety of useful constructs using a comparatively small set of primitives. %New primitives, and thus new dialects of these languages, should arise very rarely. 
%New primitives (and new \emph{dialects} of the language). %Extending a general-pandurpose language with new primitives, i.e.  forming an extended \emph{dialect} of the language, should, given a sufficiently expressive such language, be exceedingly rare. %Indeed, a diversity of dialects around a language can be taken as evidence against claims of  generality.%, this can be taken as evidence against claims of its generality. %Not all useful abstractions can be  realized, when considered comprehensively, as libraries.% The problem of achieving both simplicity and generality simultaneously with a single system remains unsolved. 
%Many languages claim to achieve this ideal. 
%For example, 
%If we measure the generality of a language by how frequently  dialects of the language emerge, 
%all major contemporary languages seem to leave room for improvement. 
%General-purpose programming languages are a dime a. 
%Among the chief goals in the design of a general-purpose programming languages must achieve \emph{stability}: they  very rarely needs to be extended or forked into dialects by the user community surrounding it. Instead, its users are able to modularly  express a wide variety of useful constructs in libraries, using the comparatively small set of primitives that the language builds in. 
Functional programming language designers have long studied small typed lambda calculi to develop a principled understanding of language-theoretic issues, like type safety, and to examine the mathematical character of language primitives of interest in isolation. These studies have then informed the design of ``full-scale''\footnote{Throughout this work, words and phrases that should be read as having an intuitive or informal meaning, rather than a strict mathematical meaning, will be introduced with quotation marks.} functional languages, which combine several such primitives and also introduce various generalizations and primitive ``embellishments''  motivated by a consideration of human factors. 
For example, major functional languages like Standard ML (SML) \cite{mthm97-for-dart,harper1997programming}, OCaml \cite{ocaml-manual} and Haskell \cite{jones2003haskell} all primitively build in record types, generalizing the nullary and binary product types that suffice in simpler calculi, because explicitly labeled components are cognitively useful to human programmers. Similarly, these languages build in derived syntactic forms (colloquially, ``syntactic sugar'') that decrease the syntactic cost of working with common library constructs, like lists.

@@

Although some of these dialects extend the type structure of the language that they are based on, the most common situation, and the one that will be our focus in this work, is that the existing semantic primitives suffice and it is only the \emph{syntactic cost} of expressing one or more constructs of interest using these primitives  that is too high. In response, library providers construct \emph{syntactic dialects} -- dialects that introduce only new derived syntactic forms. 
%Put another way, syntactic dialects can be specified by a context-independent expansion to the existing language that they are based on. 
For example, Ur/Web is a syntactic dialect of Ur (a language that itself descends from ML \cite{conf/pldi/Chlipala10}) that builds in derived forms for SQL queries, HTML elements and other datatypes used in the domain of web programming \cite{conf/popl/Chlipala15}. %Syntactic cost is often assessed qualitatively \cite{green1996usability}, though quantitative metrics can be defined. 
This is not an isolated example -- there are many other types of data that could similarly benefit from the availability of specialized derived forms (we will consider another example, regular expression patterns, in Sec. \ref{sec:syntax-examples}). 
Tools like Camlp4 \cite{ocaml-manual}, Sugar* \cite{erdweg2011sugarj,erdweg2013framework} and Racket \cite{Flatt:2012:CLR:2063176.2063195}, which we will discuss in Sec. \ref{sec:syntax-existing-approaches}, have lowered the engineering costs of constructing syntactic dialects in such situations, further contributing to their proliferation.

@@

\inferrule[T-U-abs]{
	\Delta \vdash \tau_1~\mathtt{type}\\
	\Delta~\Gamma, x:\tau \vdash \hat{e} \leadsto e : \tau'
}{
	\Delta~\Gamma \vdash \lambda x{:}\tau.\hat{e} \leadsto \lambda x{:}\tau.e : \tau \rightharpoonup \tau'
}
